---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

### set working directory, load required libraries
```{r echo=TRUE}
setwd("~/Documents/Coursera/Reproducible Research/RepData_PeerAssessment1")
library(plyr)
```

### Loading and preprocessing the data
```{r echo=TRUE}
activity <- read.csv("activity.csv", header = TRUE)
```

## What is mean total number of steps taken per day?

```{r echo=TRUE}
stepspd <- ddply(activity, .(date), summarize, steps=sum(steps))
stepspd <- subset(stepspd, !is.na(steps))
meansteps <- as.integer(mean(stepspd$steps))
mediansteps <- median(stepspd$steps)

mplot <- t(as.matrix(stepspd[-1]))
colnames(mplot) <- stepspd$date

barplot(mplot, main="Total Steps Per Day", xlab="Date", ylab="Steps")
```

The mean total steps per day is `r meansteps`.  The median total steps per day is `r mediansteps`.

## What is the average daily activity pattern?

stepstimeseries <- ts(activity, frequency=288, start=c(0,5))
plot.ts(stepstimeseries)


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
